---
title: "MSDS606_Week1_Discussion"
output: pdf_document
date: "2022-10-20"
author: "Adam Peetz"
---

```{r setup and libraries}
#set working directory
setwd("C:\\Users\\adamg\\Documents\\MSDS_660\\Week_1")
#always use session -> restart r to reboot r session prior to engagement
#load libraries
library(tidyverse)
library(scales)
```
```{r load dataset}
#load csv data
waste<-read_csv("plastic_waste.csv")
```

# Dataset desctiption

The plastic_waste.csv dataset is a clean data file. It contains 192 rows of data spread across7 columns of mixed categorical and numerical features. These features have no missing values. <br>

There was an issue with the column labels in my data set, a return, indicated by backslash n, had been added to the feature names for mismanaged_waste_2010 and mismanaged_waste_2025. I manually removed the return from the headers in the .csv file in excel and resaved it to make it easier to manipulate in R. I have attached the .csv file I modified for use with my notebook. 
```{r}
#basic dataset queries
names(waste)
ncol(waste)
length(waste)
dim(waste)
nrow(waste)

#get statistical summary of gapminder
summary(waste)
```

# Correlation of population and tons of waste per person


```{r}
#create scatter plots of population and waste generation variables
#plotcode
plot(log(population) ~ waste_gen, waste, 
     main = "Population vs Waste Generated",
     xlab="Waste Generated per Person",
     ylab="Population (logrithmic scale)")
```


